---
title: "Z0: YEScoder arrays, pixels, and images"
author: "Vincent J. Carey, stvjc at channing.harvard.edu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Z0: YEScoder arrays, pixels, and images}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
---

# R's matrix datatype

```{r lkmat}
mymat = matrix(1:25, ncol=5)
mymat
class(mymat)
mode(mymat)
dim(mymat)
```

## Get an element
```{r lk2}
mymat[3,3]
```

## Set an element
```{r lk3}
mymat[4,4] = 1234
mymat
```

## Change all elements
```{r lk5}
mymat = 5*mymat
mymat
```

# Matrix elements as "pixels"

We are speaking figuratively here.  We will
not be operating directly on your screen's pixels.
But we will get the concept of a "rasterized image" --
have a look at the wikipedia page for [raster](https://en.wikipedia.org/wiki/Raster_graphics).

```{r rebu, fig.width=4, fig.height=4}
mymat = matrix(1:25, ncol=5)
image(mymat, axes=FALSE)
```

That's a "chunky" rasterization of data in a 5x5 grid.
Let's do something more fine-grained:

```{r rebu2, fig.width=4, fig.height=4}
mymat = matrix(1:625, ncol=25)
image(mymat, col=rep(rainbow(12), 625/12), breaks=1:625, axes=FALSE)
```

We can't easily distinguish 625 colors, so there's a reuse
of colors from row to row.  But you can probably make out 625 different
little square positions in this pixelated display.

# Key distinctions: image to device

- An **image**, interpretable by the human visual system and
human reasoning, that we want to present.  In this case there
really isn't one, we have just produced an ordered sequence of
numbers in a particular computational format (R matrix).
More commonly the image will be camera-captured and formatted
into a file using the JPEG or TIFF conventions.
- **Numerical data** "representing" the image.  Conventions
are needed to structure and scale this data.
- A **rendering" process** that transmits the numerical data
to a plotting or projection system for visualization.
- A **graphics device** that receives the rendering and presents
the image to the viewer.

# Palettes manage colors

Part of the rendering process involves mapping numerical data
to visible colors.  This is done in R using "palettes".
There is a fair amount of technicality involved.  Let's scratch
the surface.

The `rainbow` function produces a palette.  Let's use it
and then explain it.

```{r myr}
plot(x=1:10, y=1:10, col=rainbow(10), pch=19, cex=5,
  xlab="x", ylab="y")
```

We see that the values plotted at coordinates (1,1), (2,2) and
so forth have different colors.  What are these colors?
Note that R actually has a color vocabulary.  Here's a little
bit of it:
```{r lkpa}
colors()[1:10]
```

But these nice terms for paint colors are not used for
the rainbow palette.
```{r lkra}
rainbow(10)
```

We can start interpreting this.  

- We have to know a little bit about hexadecimal numbers.
This is the mapping `0 -> 0, 1 -> 1, ..., 9 -> 9, 10 -> A, 11 -> B, ...,
15 -> F`.
- We have to know that
we are using the [RGB color model](https://en.wikipedia.org/wiki/RGB_color_model) by default.
In this model each color is defined by three numbers.  In R's usage,
each color intensity is a number from 0 to 255.
- We then look at the rainbow() outputs as triples. The
layout is `#abcdef`
and `ab` is the hexadecimal number for the red intensity ,
`cd` is the hex for green and `ef` is hex for blue.
- The "two digit" hex numbers `ab` are interpreted in base 10 as
`16*a + b`.  Similarly for `cd` and `ef`.

Here's a little code that decodes the two-letter hex strings.
```{r dohex}
hexi = function(str) {
 stopifnot(nchar(str) == 2)
 ss = strsplit(str, "")[[1]]
 decode = 0:15
 names(decode) = c(0:9, "A", "B", "C", "D", "E", "F")
 unname(16*decode[ss[1]] + decode[ss[2]])
}
hexi("FF")
hexi("CC")
hexi("00")
hexi("01")
```

So we can see that the rainbow output
```{r lkrabbb}
rainbow(10)
```
starts with (255,0,0), so pure red, no green
or blue.  The fourth point is `r rainbow(10)[4]`,
strong score for green, but a little red too.  This
is because the rainbow palette makes smooth transitions
between primary colors.


Exercise.  How would you change
```
plot(x=1:10, y=1:10, col=rainbow(10), pch=19, cex=5,
  xlab="x", ylab="y")
```
to illustrate a rainbow color set with 20 entries?

Try it -- do all the colors look different?


Exercise.  Substitute `topo.colors` for `rainbow` in
these plot commands.  Can you justify the color scheme
for `topo.colors`?

Hint.  Think of the display for
```
filled.contour(volcano, color.palette=topo.colors)
```
as a topographic map of a volcanic island.
