---
title: "Z1: YEScoder ImageArray introduction"
author: "Vincent J. Carey, stvjc at channing.harvard.edu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Z1: YEScoder ImageArray introduction}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
---

# ImageArrays

First we set up our R session to include a package
with some images.  A dataset called `ciftrain1k`
holds 1000 digital images that have been labeled.

```{r setup,message=FALSE}
library(YEScoder)
library(littleDeep)
library(jpeg)
data(ciftrain1k)
ciftrain1k
```

We have a special function that "plots" a handful
of the images.

```{r lkprev, echo=FALSE}
preview(ciftrain1k)
```

We'll look more closely at the collection of images later.
For now note that we can evaluate the 'variable' `ciftrain1k`

```{r lkcif}
ciftrain1k
```

and learn about the labels applied to various images:

```{r lklabs}
getTypes(ciftrain1k)[1:9]
head(table(getTypes(ciftrain1k)))
```

# Focus on one image

We can get a single image by using `[]`
with `ciftrain1k`, and extracting
the numerical data as an array.
```{r doone}
d1 = getArray(ciftrain1k[1])
dim(d1)
```

A little more manipulation allows
us to visualize the possum image directly.

```{r lkdat}
d11 = d1[1,,,,drop=TRUE]
writeJPEG(d11/255, "possum.jpg")
show_jpg("possum.jpg")
```

# The numbers beneath the image

We get familiar with arrays of numbers
when using spreadsheets like excel.
The variable `d11` is a numerical array
made of three 32 x 32 arrays stacked
on top of each other.  The first array
gives red intensity, the other two give
green and blue respectively.

We can look at the numbers using `[]`.

```{r lnum}
d11[1:3,1:3,]
```

There are 3 3x3 arrays shown here.  

We can get an appreciation for the effect
of color balance by setting all the array
values to a common value.  For example

```{r makenew}
d12 = d11
d12[,,3] = d12[,,2] # reset blue to green level
d12[,,1] = d12[,,2] # reset red to green level
d12[1:3,1:3,]
```

Now there are no color distinctions:

```{r lkgr}
writeJPEG(d12/255, "poss2.jpg")
show_jpg("poss2.jpg")
```

# Exercises

The following code produces the possum image.
Change one character so that it produces the raccoon
image.

```{r doposs, eval=FALSE}
d1 = getArray(ciftrain1k[1])
d11 = d1[1,,,,drop=TRUE]
writeJPEG(d11/255, "img.jpg")
show_jpg("img.jpg")
```

The possum and raccoon images have very different color
schemes.  We can see the relative abundances of red, green
and blue pixels in the possum image using the following:

```{r lkdens}
d1 = getArray(ciftrain1k[1])
d11 = d1[1,,,,drop=TRUE]
plot(density(d11[,,1]), lwd=2, col="red")
lines(density(d11[,,2]), lwd=2, col="green")
lines(density(d11[,,3]), lwd=2, col="blue")
```

Produce a similar display for the raccoon image.

What would you expect to see for the orange image?

