---
title: "Z2: YEScoder color concepts and modifications"
author: "Vincent J. Carey, stvjc at channing.harvard.edu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Z2: YEScoder color concepts and modifications}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
---

# RGB settings

```{r setup,message=FALSE}
library(YEScoder)
colordot(1,0,0)
colordot(1,1,0)
```

What is the color name for `colordot(.9,.5,0)`?

# An interactive approach

Start the colorapp shiny app by calling the function `colorapp()`.

Whatis a slider setting for "purple"?

Here is the code, nicely formatted:
```{r dodump,echo=FALSE,message=FALSE,results="hide"}
appcode = dumpapp()
```
```{r showit,echo=FALSE}
cat(appcode, sep="\n")
```

It is a function with no arguments that defines two objects:

- ui: the user interface, built using shiny components that
compose html, css and javascript to define panels, tabs, and
sliders
- server: a function with arguments input and output (and
potentially session) that defines how user activities lead
to computations that are transformed and passed back to the
browser

When those objects are properly defined, the shiny package
function `runApp` will operate on them to produce an interactive
app in the browser.  The app has access to data accessible in
the R session in which the app was started.

Look at the code defining the "orange" tab.  What is being done
to the color intensities on the basis of slider settings?
